---
title: "Class 12:Pathway Analysis from RNA-Seq Results"
author: 'Hong (PID: A16558957)'
date: "2/26/2022"
output: pdf_document
---

```{r}
library(DESeq2)
```


```{r}
metaFile <- "GSE37704_metadata.csv"
countFile <-"GSE37704_featurecounts.csv"
```

```{r}
countData = read.csv(countFile, row.names=1)
head(countData)
```

```{r}
colData = read.csv(metaFile, row.names=1)
head(colData)
```

>Q. Complete the code below to remove the troublesome first column from countData

```{r}
countData <- as.matrix(countData[,-1])
head(countData)
```

> Complete the code below to filter countData to exclude genes (i.e. rows) where we have 0 read count across all samples (i.e. columns).

```{r}
zero.vals <- which(countData[,1:2]==0, arr.ind=TRUE) 
to.rm <- unique(zero.vals[,1])
mycounts <- countData[-to.rm,]
head(mycounts)
```

```{r}
dds = DESeqDataSetFromMatrix(countData=mycounts,
                             colData=colData,
                             design=~condition)
dds = DESeq(dds)

dds
```

```{r}
res = results(dds, contrast=c("condition", "hoxa1_kd", "control_sirna"))
```

>Call the summary() function on your results to get a sense of how many genes are up or down-regulated at the default 0.1 p-value cutoff.

```{r}
summary(res)
```

```{r}
plot( res$log2FoldChange, -log(res$padj) )
```

>Q. Improve this plot by completing the below code, which adds color and axis labels

```{r}
mycols <- rep("gray", nrow(res) )
mycols[ abs(res$log2FoldChange) > 2 ] <- "red"
inds <- (res$padj < 0.01) & (abs(res$log2FoldChange) > 2 )
mycols[ inds ] <- "blue"
plot( res$log2FoldChange, -log(res$padj), col=mycols, xlab="Log2(FoldChange)", ylab="-Log(P-value)" )
```

>Q. Use the mapIDs() function multiple times to add SYMBOL, ENTREZID and GENENAME annotation to our results by completing the code below.

```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")

columns(org.Hs.eg.db)
```

```{r}
res$symbol = mapIds(org.Hs.eg.db,
                    keys=row.names(res), 
                    keytype="ENSEMBL",
                    column="SYMBOL",
                    multiVals="first")

res$entrez = mapIds(org.Hs.eg.db,
                    keys=row.names(res),
                    keytype="ENSEMBL",
                    column="ENTREZID",
                    multiVals="first")

res$name =   mapIds(org.Hs.eg.db,
                    keys=row.names(res),
                    keytype="ENSEMBL",
                    column="GENENAME",
                    multiVals="first")

head(res, 10)
```

>Q. Finally for this section let's reorder these results by adjusted p-value and save them to a CSV file in your current project directory.

```{r}
res = res[order(res$pvalue),]
write.csv(res, file ="deseq_results.csv")
```



```{r}
library(pathview)
```

```{r}
library(gage)
library(gageData)

data(kegg.sets.hs)
data(sigmet.idx.hs)
head(kegg.sets.hs, 3)
```

```{r}
foldchanges = res$log2FoldChange
names(foldchanges) = res$entrez
head(foldchanges)
```

```{r}
keggres = gage(foldchanges, gsets=kegg.sets.hs)
attributes(keggres)
```

```{r}
head(keggres$less)
```

```{r}
pathview(gene.data=foldchanges, pathway.id="hsa04110")
```

```{r}
pathview(gene.data=foldchanges, pathway.id="hsa04110", kegg.native=FALSE)
```

```{r}
keggrespathways <- rownames(keggres$greater)[1:5]

# Extract the 8 character long IDs part of each string
keggresids = substr(keggrespathways, start=1, stop=8)
keggresids
```

```{r}
pathview(gene.data=foldchanges, pathway.id=keggresids, species="hsa")
```

>Q. Can you do the same procedure as above to plot the pathview figures for the top 5 down-reguled pathways?

```{r}
keggrespathwaysdown <- rownames(keggres$less)[1:5]

# Extract the 8 character long IDs part of each string
keggresidsdown = substr(keggrespathwaysdown, start=1, stop=8)
keggresidsdown

pathview(gene.data=foldchanges, pathway.id=keggresidsdown, species="hsa")
```

```{r}
data(go.sets.hs)
data(go.subs.hs)

# Focus on Biological Process subset of GO
gobpsets = go.sets.hs[go.subs.hs$BP]

gobpres = gage(foldchanges, gsets=gobpsets, same.dir=TRUE)

lapply(gobpres, head)
```

```{r}
sig_genes <- res[res$padj <= 0.05 & !is.na(res$padj), "symbol"]
print(paste("Total number of significant genes:", length(sig_genes)))
```

```{r}
write.table(sig_genes, file="significant_genes.txt", row.names=FALSE, col.names=FALSE, quote=FALSE)
```

>Q: What pathway has the most significant “Entities p-value”? Do the most significant pathways listed match your previous KEGG results? What factors could cause differences between the two methods?

