---
title: 'Class 6: R Functions'
author: 'Hong (PID: A16558957)'
date: "2/3/2022"
output: pdf_document
---

Today we will explore R functions.

We will start with calculating a grade for these example students.

```{r}
# Example input vectors to start with
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)
```

We could use the mean() function to calculate an average.

```{r}
mean(student2, na.rm = TRUE)
```

How does the is.na() function work? Let's try it out on student2.

```{r}
x <- is.na(student2)
```

We can use this result to get at our NA values (i.e, the TRUE positions).

```{r}
student2[ is.na(student2)] <- 0
student2
```
```{r}
student3[is.na(student3)] <- 0
student3
mean(student3)
```

```{r}
student2[is.na(student2)] <- 0
mean(student2[-which.min(student2)])
```
```{r}
student3[is.na(student3)] <- 0
mean(student3[-which.min(student3)])
```


```{r}
student1[-which.min(student1)]
```


```{r}
x <- student3
x[is.na(x)] <- 0
x <- x[-which.min(x)]
mean(x)
```

We are close to our working code snipet that willl be the body of our first function.

```{r}
# First set NA values to zero
x[is.na(x)] <- 0
# Remove lowest score and calculate average
mean(x[-which.min(x)])

```

Now we can turn this into our first function. We will call this function 'grade'

All R functions have 3 things 
- a name
- input arguments
- body
```{r}
grade <- function(x){
# First set NA values to zero
  x[is.na(x)] <- 0
# Remove lowest score and calculate average
  mean(x[-which.min(x)])
}
```

```{r}
grade(student3)
```

```{r}
gradebook <- read.csv("https://tinyurl.com/gradeinput", row.names = 1)
head(gradebook)
```

We can use the 'apply()' function to grade the whole class.

```{r}
scores <- apply(gradebook, 1, grade)
scores
```
> Q2. Who is the top scoring student overall in the gradebook?

```{r}
which.max(scores)
```
> Q3. which homework was toughest on students (i.e. obtained the lowest scores overall?

```{r}
hw.mean <- (apply(gradebook,2,mean, na.rm = TRUE))
which.min(hw.mean)
```
```{r}
boxplot(gradebook)
```




