---
title: "Class 8: Mini Project"
author: 'Hong (PID: A16558957)'
date: "2/10/2022"
output: pdf_document
---

Unsupervised Learning Analysis of Human Breast Cancer Cells

Here we read data from the University of Wisconsin Medical Center on breast cancer patients

```{r}
fna.data <- "WisconsinCancer.csv"
wisc.df <- read.csv(fna.data, row.names=1)
wisc.df
```

```{r}
wisc.data <- wisc.df[,-1]
wisc.data
```

```{r}
diagnosis <- as.factor(wisc.df$diagnosis)
```

>Q1. How many observations are in this dataset?

How many rows (i.e. patients)
```{r}
nrow(wisc.df)
```

How many columns (i.e. variables)
```{r}
ncol(wisc.df)
```

>Q2. How many of the observations have a malignant diagnosis?

```{r}
table(diagnosis)
```

>Q3. How many variables/features in the data are suffixed with _mean?

First I need to get at where the column names are stored
```{r}
a <- (grep("_mean", colnames(wisc.df)))
length(a)
```

```{r}
colMeans(wisc.data)

apply(wisc.data,2,sd)
```

## Principal Component Analysis (PCA)

Here we need to scale the data before PCA as the various variables (i.e. columns) have very different scales.

```{r}
wisc.pr <- prcomp( wisc.data, scale = TRUE)
summary(wisc.pr)
```

Now I will make my mian result: the "PCA plot" (a.k.a. "score plot", PC1 vs PC2 plot)
```{r}
plot(wisc.pr$x[,1:2], col = diagnosis)
```

>Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?

.4427

>Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?

3

>Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?

7

```{r}
biplot(wisc.pr)
```

>Q7. What stands out to you about this plot? Is it easy or difficult to understand? Why?

Black center circle with a lot of red lines. It is very difficult to understand, because it is really messy and the labels and overlapping each other.


>Q8. Generate a similar plot for principal components 1 and 3. What do you notice about these plots?

The red dots are more gathered comparing to the plot of PC1 vs PC2, which is not as clean as the first plot.
```{r}
plot(wisc.pr$x[,c(1,3)], col = diagnosis)
```

```{r}
df <- as.data.frame(wisc.pr$x)
df$diagnosis <- diagnosis
library(ggplot2)
ggplot(df) + 
  aes(PC1, PC2, col=diagnosis) + 
  geom_point()
```



>Q9. For the first principal component, what is the component of the loading vector (i.e. wisc.pr$rotation[,1]) for the feature concave.points_mean?

 -0.26085376
```{r}
wisc.pr$rotation[,1]
```


>Q10. What is the minimum number of principal components required to explain 80% of the variance of the data?

5 PCs

##Hierarchical Clustering

First let's try clustering the raw data. This does not look promising...
```{r}
hc <- hclust( dist(wisc.data))
plot(hc)
```

```{r}
data.scaled <- scale(wisc.data)
data.dist <- dist(data.scaled)
wisc.hclust <- hclust(data.dist, method = "complete")
```

```{r}
plot(wisc.hclust)
abline(h = 19, col="red", lty=2)
```

>Q11. Using the plot() and abline() functions, what is the height at which the clustering model has 4 clusters?

19

```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k= 4)
table(wisc.hclust.clusters, diagnosis)
```

>Q12. Can you find a better cluster vs diagnoses match by cutting into a different number of clusters between 2 and 10?

clusters of 2 is probably the best, because the lower the number the better the result. When it is cut into 2 clusters, the result is better and less distributed compared to 10 clusters.
```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k= 2)
table(wisc.hclust.clusters, diagnosis)
```


>Q13. Which method gives your favorite results for the same data.dist dataset? Explain your reasoning.

I think complete method gives my favorite result since the height is not too long like ward.2D, it doesn't cut the top like average, and it has a better graph than single method.

We can combine methods to be useful. We can take our PCA results and apply clustering to them.

Here we will take the first 3 PCs for clustering
```{r}
pcdist <- dist(wisc.pr$x[,1:3])
wisc.pr.hclust <- hclust(pcdist, method = "ward.D2")
plot(wisc.pr.hclust)
```

This looks much more promising than our previous

```{r}
grps <- cutree(wisc.pr.hclust, k=2)
```

```{r}
plot(wisc.pr$x[,1:2] , col = grps)
```

How well do my clusters agree with the expert M/B values
```{r}
table(diagnosis,grps)
```

```{r}
adist <- dist(wisc.pr$x[,1:7])
wisc.pr.hclust <- hclust(adist, method="ward.D2")
wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k=2)
table(wisc.pr.hclust.clusters,diagnosis)
```
>Q15. How well does the newly created model with four clusters separate out the two diagnoses?

It works pretty well since they are fairly distributed.

>Q16. How well do the k-means and hierarchical clustering models you created in previous sections (i.e. before PCA) do in terms of separating the diagnoses? Again, use the table() function to compare the output of each model (wisc.km$cluster and wisc.hclust.clusters) with the vector containing the actual diagnoses.

It works pretty well since they are fairly distributed.

>Q17. Which of your analysis procedures resulted in a clustering model with the best specificity? How about sensitivity?

I think hierarchical clustering would be the best specificity and PCA would be the best sensitivity.

```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

```{r}
plot(wisc.pr$x[,1:2], col= diagnosis)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```

>Q18. Which of these new patients should we prioritize for follow up based on your results?

We should prioritize the blue cirlce patients.



