---
title: "Class 14:COVID-19 Vaccination Rates"
author: 'Hong (PID: A16558957)'
date: "3/3/2022"
output: 
  pdf_document : 
    latex_engine: xelatex

---

```{r}
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```

>Q1. What column details the total number of people fully vaccinated? 

persons_fully_vaccinated

>Q2. What column details the Zip code tabulation area? 

zip_code_tabulation_area

>Q3. What is the earliest date in this dataset? 

2021-01-05

>Q4. What is the latest date in this dataset? 

2022-03-01

```{r}
vax$as_of_date[nrow(vax)]
```

```{r}
skimr::skim(vax)
```

>Q5. How many numeric columns are in this dataset? 

9

>Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column? 

18174

>Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)? 

0.17

```{r}
library(lubridate)
```

```{r}
today()
```

First I have to make sure my covid vaccination data column is in lubridate format

```{r}
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
today() - vax$as_of_date[1]
```

```{r}
today()-vax$as_of_date[nrow(vax)]
```

> Q9. How many days have passed since the last update of the dataset? 

2

> Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)? 

```{r}
length(unique(vax$as_of_date))
```

## Focus on the San Diego area

```{r}
library(zipcodeR)
```

```{r}
geocode_zip('92037')
```

```{r}
zip_distance('92037','92109')
```

```{r}
reverse_zipcode(c('92037', "92109") )
```

```{r}
vax$county == "San Diego"
```

An oftern more convenient way to do this type of "filtering" (a.k.a. subsetting) is with the **dplyr**.

```{r}
library(dplyr)

sd <- filter(vax, county == "San Diego")

dim(sd)
```

```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
nrow(sd.10)
```

> Q11. How many distinct zip codes are listed for San Diego County? 

```{r}
length(unique(sd$zip_code_tabulation_area))
```


> Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset? 

92154

```{r}
sd[which.max(sd$age12_plus_population),"zip_code_tabulation_area"]
```


Using dplyr select all San Diego “county” entries on “as_of_date” “2022-02-22” and use this for the following questions.

Let's do this with the most recent date ub tge data-set (2022-03-01)

> Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2022-02-22”? 

```{r}
sd.latest <- filter(sd, as_of_date == "2022-03-01" )
summary(sd.latest$percent_of_population_fully_vaccinated, na.rm = TRUE)
```


> Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2022-02-22”?

```{r}
hist(sd.latest$percent_of_population_fully_vaccinated, breaks = 30)
```

```{r}
library(ggplot2)

ggplot(sd.latest) + 
  aes(percent_of_population_fully_vaccinated) +
  geom_histogram()
```

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population

```

>Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
baseplot <- ggplot(ucsd) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  theme_bw() +
  ylim(c(0,1)) +
  labs(x = "Date", y="Percent Vaccinated")

baseplot
```

> Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-02-22”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
baseplot + 
  labs(title = "Vaccination Rate for CA 92037(UCSD)")
```

Add a line showing the average vaccination rate for all zip code areas with a population just as large as 92037

```{r}
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2022-03-01")
head(vax.36)
```

```{r}
line <- mean(vax.36$percent_of_population_fully_vaccinated, na.rm = T)
line
```

```{r}
baseplot + geom_hline(yintercept = line, color = "red", linetype = 2)
```


> Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-02-22”?

```{r}
summary(vax.36$percent_of_population_fully_vaccinated, na.rm = TRUE)
```


> Q18. Using ggplot generate a histogram of this data.

```{r}
ggplot(vax.36) +
  aes(percent_of_population_fully_vaccinated) +
  geom_histogram() + 
  xlim(c(0,1))
```


> Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
vax %>% filter(as_of_date == "2022-03-01") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
```

```{r}
vax %>% filter(as_of_date == "2022-03-01") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)
```

The 92109 and 92040 ZIP code areas are below the average value calculated above.

> Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)


ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccination",
       title="Vaccination rate across California",
       subtitle=" Only areas with a population above 36k are shown") +
  geom_hline(yintercept = line, linetype=2)

```


> Q21. How do you feel about traveling for Spring Break and meeting for in-person class afterwards?

I don't think traveling for Spring Break is going to be too fun, because we would have like less than a week to rest. So I would rather sleeping than traveling in Spring Break. I feel normal if the classes are going to be in-person.






