---
title: 'Class 10: Genome Informatics'
author: 'Hong (PID: A16558957)'
date: "2/17/2022"
output: pdf_document
---

# Examine 1000 Genome Data

>Q5: What proportion of the Mexican Ancestry in Los Angeles sample population (MXL) are
homozygous for the asthma associated SNP (G|G)?

14.06

```{r}
# Read genotype file from Ensemble
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
round(table(mxl$Genotype..forward.strand.)/nrow(mxl)*100,2)
```

What about a different population? Here we take the British in England and Scotland (GBR)

```{r}
gbr <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
round(table(gbr$Genotype..forward.strand.)/nrow(gbr) * 100,2)
```

## Expression by Genotype analysis

I want read my RNA-Seq expression results into R. This file is not a CSV but rather has fields separated by space.

```{r}
x <- read.table("rs8067378_ENSG00000172057.6.txt")
x
```

First try at this question. Is the mean expression different based on genotype?

```{r}
x$geno == "G/G"
```

```{r}
summary(x[x$geno =="G/G",]$exp)
```

Now we will look at the other genotypes

```{r}
table(x$geno)
```

```{r}
summary(x[x$geno == "A/A",3])
summary(x[x$geno == "A/G",3])
summary(x[x$geno == "G/G",3])
```

## Make a summary overview figure

Make a boxplot figure...
```{r}
library(ggplot2)

ggplot(x) + aes(geno, exp, fill = geno) + geom_boxplot(notch = TRUE)
```






